Hierarchical Time Series Forecasting in Emergency Medical Services

### 2023-07-22 journal_correspondence/response_letter/response_letter_JSR.qmd

*I liked your point about reproducibility very much. It is great that you will released you code on Github. Would it be possible to release your data as a benchmark data set for the research community? I understand that there may be NDAs involved, but publicly available data sets help the field advance.*

*This is minor, but there seem to be issues with your bibliography software. Sometimes it inserts initials in the text citations, other times not. Sometimes commas are missing, e.g., "Haugsbø Hermansen and Mengshoel (2021)". There are other writing errors, e.g., sometimes "neural network" is capitalized and other times not. It is not a proper noun.*

Response:  Thank you very much for considering our paper for publication. We are sincerely grateful to the reviewers for providing such valuable feedback, and we highly value the comments raised by all reviewers and the Associate Editor. We have devoted significant effort to carefully reflect on the concerns raised and have made every effort to address them in the revised version of the manuscript.

We agree that the reviewers have raised important questions and provided valuable recommendations regarding the finer details of our study, and we have taken all of these into account in the revised manuscript. In response to the suggestion to look beyond the point estimate, we have incorporated an example that highlights the advantages of doing so. Furthermore, we have thoroughly addressed the specific suggestions provided by Reviewers 1 and 2 in the revised manuscript. Additionally, we have conducted further analysses and responded to Reviewer 3's remarks concerning the performance of the Naïve model (name changed to Stationary in the revised version) and Figure 4. To emphasize the significance of our study, we have included additional paragraphs in the paper that highlight its importance and potential extensions. Specifically, in response to R3's remark, we have discussed how the framework can be extended to incorporate the decision-making process and assess the implications of forecasts on crucial aspects like cost allocation and resource management, which are vital to planners. Finally, we elaborate on how our current study has broader implications and can be generalized to other areas within the service industries. This highlights the wide applicability and potential impact of our research in various contexts.

Our manuscript is written using [Quarto](https://quarto.org/), an open-source scientific and technical publishing system designed for publishing reproducible and production-quality articles. To ensure transparency and accessibility, we have established a GitHub repository containing all the necessary files, analyses, R codes, and related materials used in creating the article. Currently, the repository link remains private to avoid revealing the authors' identities in the paper. However, upon request, we are prepared to share this link with the Editor, Associate Editor, and all reviewers. Once the paper is accepted for publication, we will include the repository link in the article to facilitate further exploration and validation of our work.

@@

_All reviewers have questions and recommendations regarding the details of your time series. They seem to be justified and not in conflict with each other. So, consider implementing them._

_All reviewers have questions regarding your look “beyond the point estimate”. Consider them. It might help to also come up with an example that outlines the advantage of going beyond point estimates in greater detail._

_R1 has a long list of precise and helpful suggestions for improving the manuscript. It looks like they are all justified, and it should be manageable to implement them. The same holds for R2 and R3. R3’s remark regarding Figure 4 and the performance of the naïve model that seems to be not much worse is particularly crucial._

**Response:** Thank you for your feedback and for summarizing the main points from the reviewers. We appreciate your suggestion to highlight the insights provided by the forecasting exercise, and we have worked on the manuscript to make this more explicit in the revised manuscript. We agree that the reviewers have raised important questions and recommendations regarding the details of our study, and we have addressed these in the revised manuscript. Regarding the recommendation to look beyond the point estimate, we have taken the reviewers' suggestions into account and provide an example that illustrates the advantage of doing so. We have responded and implemented suggestion provided by R1 and R2 in the revised manuscript. We have also conducted further analyses to respond to R3's remark regarding the performance of the Naïve model (name changed to Stationary in the revised version) and Figure 4. We have also included additional paragraphs in the paper to highlight the importance of the current study and how it can be extended to incorporate the decision making process and evaluate the implication of forecasts on decision impacts and utilities important to planners such as cost, allocating resources, etc. Finally, we added a paragraph to the end of the conclusion section to highlight how the current study can be generalized to other areas in service industries.

Our manuscript is written using [Quarto](https://quarto.org/), an open-source scientific and technical publishing system designed for publishing reproducible and production-quality articles. To ensure transparency and accessibility, we have established a GitHub repository containing all the necessary files, analyses, R codes, and related materials used in creating the article. Currently, the repository link remains private to avoid revealing the authors' identities in the paper. However, upon request, we are prepared to share this link with the Editor, Associate Editor, and all reviewers. Once the paper is accepted for publication, we will include the repository link in the article to facilitate further exploration and validation of our work.

### 2023-07-22 manuscript/02-literature.qmd
resubmit word document

We note a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research. These limitations are summarized as following:

@@

```{r summarylit, echo=FALSE, cache=FALSE}
library(kableExtra)
li <- readxl::read_xlsx("../literature_table/table.xlsx")
knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies in forecasting FOR AMBULANCE SERVICES") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |>  kableExtra::landscape()
```

>>>>

@tbl-literature provides a summary of some studies in the literature on forecasting in Emergency Medical Services. We note a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research. These limitations are summarized as following:

@@

This paper concerns the problem of hierarchical forecasting in EMS and generates and evaluates both point and probabilistic forecast across different levels of the hierarchy, hence addressing some important gaps identified in the literature.

```{r summarylit, echo=FALSE, cache=FALSE}
#| label: tbl-literature
library(kableExtra)
li <- readxl::read_xlsx("../literature_table/table.xlsx")
knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies on forecasting in Emergency Medical Services") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |>  kableExtra::landscape()
```

### 2023-07-22 manuscript/main.qmd
add the Github repo
resubmit word document

---
title: "Hierarchical Time Series Forecasting in Emergency Medical Services"
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations:
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet. Such demand forecasts are required at national, regional, and sub-regional levels, and must take account of the nature of incidents and their priorities. These forecasts are often generated independently by different teams within the organization. As a result, forecasts at different levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service. To address this issue, we exploit the hierarchical and grouped structure of the demand time series and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation. The methods are applied to daily incident data from an ambulance service in Great Britain, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area. We use an ensemble of forecasting models and show that the resulting forecasts are better than any individual forecasting model. We validate the forecasting approach using time series cross validation.
keywords:
  - healthcare
  - emergency services
  - forecast reconciliation
  - ambulance demand
  - regression
bibliography: bibliography.bib
execute:
  echo: false
  message: false
  warning: false
  cache: true
pdf-engine: pdflatex
prefer-html: true
format:
  html:
    number-sections: true
  docx:
    number-sections: true
  elsevier-pdf:
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{todonotes,mathtools,bm,amsmath,mathpazo}
        \mathtoolsset{showonlyrefs}
        \setlength{\parindent}{0cm}
    journal:
      name: Journal of Service Research
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

>>>>

---
title: "Hierarchical Time Series Forecasting in Emergency Medical Services"
# author:
#   - name: Bahman Rostami-Tabar
#     email: rostami-tabarb@cardiff.ac.uk
#     affiliations:
#         - id: carbs
#           name: Cardiff University
#           department: Cardiff Business School
#           country: United Kingdom
#           postal-code: CF10 3EU
#     attributes:
#         corresponding: true
#   - name: Rob J. Hyndman
#     email: Rob.Hyndman@monash.edu
#     affiliations:
#         - id: monash
#           name: Monash University
#           department: Department of Econometrics and Business Statistics
#           country: Australia
#           postal-code: VIC 3800
abstract: |
  Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet. Such demand forecasts are required at national, regional, and sub-regional levels, and must take account of the nature of incidents and their priorities. These forecasts are often generated independently by different teams within the organization. As a result, forecasts at different levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service. To address this issue, we exploit the hierarchical and grouped structure of the demand time series and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation. The methods are applied to daily incident data from an ambulance service in Great Britain, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area. We use an ensemble of forecasting models and show that the resulting forecasts are better than any individual forecasting model. We validate the forecasting approach using time series cross validation.
keywords:
  - healthcare
  - emergency services
  - forecast reconciliation
  - ambulance demand
  - regression
bibliography: bibliography.bib
execute:
  echo: false
  message: false
  warning: false
  cache: true
pdf-engine: pdflatex
prefer-html: true
format:
  html:
    number-sections: true
  docx:
    number-sections: true
  elsevier-pdf:
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{todonotes,mathtools,bm,amsmath,mathpazo}
        \mathtoolsset{showonlyrefs}
        \setlength{\parindent}{0cm}
    journal:
      name: Journal of Service Research
      formatting: preprint
      model: 3p
      cite-style: authoryear
---
